Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-20 latex/article.tex
minor

In our context, we consider digital shapes (any subset of $\Z^d$) and boundaries
of digital shapes. We denote by $\DSh$ the Gauss digitization of
$\Shape$ in a $d-$dimensional grid with grid step $h$, \emph{i.e.}
$\DSh = \Shape \cap (h\Z)^d$. For such digitized set $\DigShape$,
$\Body{\DigShape}{h}$ denotes its topological boundary, seen as a cellular Cartesian
complex (See Fig. \ref{fig:notations}-b). It is thus composed of $0$-cells and
$1$-cells (resp. \emph{pointels} and \emph{linels}), and, for $d=3$, with
$2$-cells ({\em surfels}), embedded in the digital grid.

@@

We have proved that estimating the 2D curvature and 3D mean curvature on digital
objets with volumetric integral information achieve to the multigrid convergence
property (setting $R = kh^\frac{1}{3}$ on convex shapes with at least
$C^3$-boundary and bounded curvature). For short, when the digital objet become
finer and finer, \ie when the digitization step $h$ tends to zero, the estimate
quantity on $\DSh$ should converge (theoretically and experimentally) to the associated one on $\Shape$.

@@

In \cite{CVIU2014}
the 3D digital principal curvature estimators $\kappa^1_{R}$ and $\kappa^2_{R}$
on $Z\subset\Z^3$

\begin{Definition}
  %
  Given $\DigShape \subset \Z^3$ and
  $h$ a gridstep, {\em 3D digital principal curvature estimators} are defined for any pointel
  $\vp \in \Body{\DigShape}{h}$ as:
  %
  \begin{align}
    \hat{\kappa}^{1}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_2 - 3\hat{\lambda}_1)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),\label{eq:dig-curvature-estimator-k1}\\
    \hat{\kappa}^{2}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_1 - 3\hat{\lambda}_2)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $R_d = \frac{R}{h}$ is the digitized radius of the ball, and
  $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}

% When dealing with digital data in a multigrid framework, we must pay
% attention to on how quantities expend where decreasing the grid step
% $h$. For example, if you fixed $R$, the digital radius will be $R_d
% = \frac{5}{h}$ and increase when $h$ tends to zero.
Authors of \cite{CVIU2014} have proved the convergence of these
estimators when setting the ball radius dependent on the
grid step $h$: $R = kh^\frac{1}{3}$, where $k$ is a constant related
to the maximal curvature of the shape. More formally, we get:
\begin{align}
  \hat{\kappa}^1_{R}(\DigF{\Shape}{h},p) &= \kappa^1(X,x) + O(h^{\frac{1}{3}}),\\
  \hat{\kappa}^2_{R}(\DigF{\Shape}{h},p) &= \kappa^2(X,x) + O(h^{\frac{1}{3}}),
\end{align}
when the ball radius is $R = kh^\frac{1}{3}$.

>>>>

In our context, we consider digital shapes (any subset of $\Z^d$) and
boundaries of digital shapes. We denote by $\DSh$ the Gauss
digitization of $\Shape$ in a $d-$dimensional grid with grid step $h$,
\emph{i.e.}  $\DSh = \Shape \cap (h\Z)^d$. For such digitized set
$\DigShape$, $\Body{\DigShape}{h}$ \todo{DC: $\partial\Body{.}{h}$,
  pas $\Body{.}{h}$, c'est pas un bord fait de i-cell} denotes its
topological boundary, seen as a cellular Cartesian complex (See
Fig. \ref{fig:notations}-b). It is thus composed of $0$-cells and
$1$-cells (resp. \emph{pointels} and \emph{linels}), and, for $d=3$,
with $2$-cells ({\em surfels}), embedded in the digital grid.

@@

We have proved that estimating the 2D curvature and 3D mean curvature on digital
objets with volumetric integral information have the multigrid convergence
property (setting $R = kh^\frac{1}{3}$ on convex shapes with at least
$C^3$-boundary and bounded curvature) \cite{CVIU2014}. For short, when the digital objet becomes
finer and finer, \ie when the digitization step $h$ tends to zero, the estimated
quantity on $\DSh$  converges (theoretically and experimentally) to
the associated one on $\partial\Shape$. More precisely, we have:

@@

In \cite{CVIU2014}, we have also defined the 3D digital principal curvature estimators $\kappa^1_{R}$ and $\kappa^2_{R}$
on $Z\subset\Z^3$:
\begin{Definition}
  %
  Given $\DigShape \subset \Z^3$ and
  $h$ a gridstep, {\em 3D digital principal curvature estimators} are defined for any pointel
  $\vp \in \Body{\DigShape}{h}$ as:
  %
  \begin{align}
    \hat{\kappa}^{1}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_2 - 3\hat{\lambda}_1)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),\label{eq:dig-curvature-estimator-k1}\\
    \hat{\kappa}^{2}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_1 - 3\hat{\lambda}_2)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $R_d = \frac{R}{h}$ is the digitized radius of the ball, and
  $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}

% When dealing with digital data in a multigrid framework, we must pay
% attention to on how quantities expend where decreasing the grid step
% $h$. For example, if you fixed $R$, the digital radius will be $R_d
% = \frac{5}{h}$ and increase when $h$ tends to zero.
Authors of \cite{CVIU2014} have proved the convergence of these
estimators when setting the ball radius $h$ in $R = kh^\frac{1}{3}$,
where $k$ is a constant related to the maximal curvature of the
shape. More formally, we get:
\begin{align}
  \hat{\kappa}^1_{R}(\DigF{\Shape}{h},p) &= \kappa^1(X,x) + O(h^{\frac{1}{3}}),\\
  \hat{\kappa}^2_{R}(\DigF{\Shape}{h},p) &= \kappa^2(X,x) + O(h^{\frac{1}{3}}),
\end{align}
when the ball radius is $R = kh^\frac{1}{3}$.
Additionally, eigenvectors associated with $\hat{\lambda}_1$ and
$\hat{\lambda}_2$ correspond to principal curvature directions. The
smallest eigvenvector corresponds to the normal direction at $p$.
\todo[inline]{finir en citant le chapitre (ou CVIU) pour convergence directions}

\todo[inline]{petite phrase pour dire que R depend de la géométrie
  (citer DGCI pour paramter free, ou SMI pour dire qu'on peut s'en
  servir pour feature) et qu'il faut explorer R}
